Structure constants and short-time simulations of open collective quantum spin systems

### 2019-03-18 soc_squeezing/collective_paper/collective_spin.tex
updates (largely wording/symbol changes) to the collective spin paper

%%% figures
\usepackage{graphicx} % for figures
\graphicspath{{./figures/}} % set path for all figures

@@

\title{Short-time simulation of open collective quantum spin systems}

@@

Collective spin systems are a versatile resource in quantum science
for a range of applications including quantum-enhanced metrology,
quantum simulation, and quantum information.  Theoretical interest in
such systems dates back to the mid-twentieth century with the
introduction of the Dicke model\cite{dicke1954coherence} that
describes atoms cooperatively interacting with a single mode of the
radiation field and the Lipkin-Meshkov-Glick (LMG) model, a toy model
for testing many-body approximation methods in contemporary nuclear
physics\cite{lipkin1965validity, meshkov1965validity,
  glick1965validity}.  On the experimental side, the development of
advanced trapping, cooling, and control techniques in atomic,
molecular, and optical (AMO) systems have enabled the realization of
collective spin models in a broad range of platforms, including atomic
ensembles\cite{takano2009spin, appel2009mesoscopic, chen2010heralded,
  schleier-smith2010states, chen2011conditional}, Bose-Einstein
condensates\cite{riedel2010atomchipbased, gross2010nonlinear,
  klinder2015dynamical}, ultracold Fermi
gasses\cite{martin2013quantum, smale2018observation}, trapped
ions\cite{bohnet2016quantum}, optical
cavities\cite{leroux2010implementation, bohnet2014reduced,
  cox2016deterministic, hosten2016measurement, hosten2016quantum,
  zhiqiang2017nonequilibrium, lewis-swan2018robust,
  norcia2018cavitymediated}, and molecular
sensors\cite{jones2009magnetic}. These implementations have triggered
intensive investigations of rich phenomena, including studies of
quantum criticality\cite{latorre2005entanglement,
  alcalde2007functional, morrison2008dissipationdriven,
  morrison2008collective, sarandy2009classical, wang2012quantum,
  kessler2012dissipative, majd2014lmg}, non-equilibrium
phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
  bhattacherjee2014nonequilibrium, klinder2015dynamical,
  maghrebi2016nonequilibrium, zhiqiang2017nonequilibrium,
  lang2018concurrence, smale2018observation}, and precision
metrology\cite{wineland1992spin, itano1993quantum,
  kitagawa1993squeezed, ma2011quantum, takano2009spin,
  appel2009mesoscopic, chen2010heralded, schleier-smith2010states,
  chen2011conditional, riedel2010atomchipbased, gross2010nonlinear,
  zhong2010simplified, agarwal1997atomic, lau2014proposal,
  huang2015quantum, wineland1992spin, kitagawa1993squeezed,
  schleier-smith2010squeezing, leroux2010implementation,
  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
  jones2009magnetic}.

One of the primary motivations for studying collective spin systems is
their application to quantum-enhanced metrology.  Quantum projection
noise limits the error $\Delta\phi$ in the measurement of a phase
angle $\phi$ with $N$ independent spins to
$\Delta\phi\sim1/\sqrt{N}$\cite{wineland1992spin, itano1993quantum,
  ma2011quantum}.  Collective spin systems enable a means to break
through this limit via the preparation of many-body entangled states
such as NOON\cite{jones2009magnetic, chen2010heralded,
  zhong2010simplified}, spin-cat\cite{agarwal1997atomic,
  lau2014proposal, huang2015quantum}, and spin-squeezed
states\cite{wineland1992spin, kitagawa1993squeezed, takano2009spin,
  appel2009mesoscopic, schleier-smith2010states,
  schleier-smith2010squeezing, leroux2010implementation,
  riedel2010atomchipbased, gross2010nonlinear, chen2011conditional,
  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
  ma2011quantum} that allow for measurement errors
$\Delta\phi\sim1/N^\epsilon$ with $\epsilon>1/2$.  Generating such
states typically requires nonlinear dynamics that are realized using
long-range collisional, photon-, or phonon-mediated interactions.
Although a truly collective spin model requires uniform, all-to-all
interactions, as long as measurements do not distinguish between
constituent particles even non-uniform systems can be effectively
described by a uniform model with renormalized
parameters\cite{hu2015entangled}.

In the absence of decoherence, permutation symmetry and total spin
conservation divide the total Hilbert space of a collective spin
system into superseletion sectors that grow only linearly with system
size $N$, thereby admitting efficient classical simulation of its
dynamics.  Decoherence will generally violate total spin conservation
and require the use of density operators, increasing the dimension of
accessible state space to $\sim N^3$\cite{hartmann2016generalized,
  xu2013simulating} and allowing for exact simulations of
$N\lesssim100$ particles.  If decoherence is sufficiently weak,
dynamics can be numerically solvable for $N\lesssim10^5$ particles via
``quantum trajectory'' Monte Carlo methods\cite{plenio1998quantumjump,
  zhang2018montecarlo} (also known as ``quantum jump'' or ``Monte
Carlo wavefunction'' methods) that can reproduce all expectation
values of interest.  When decoherence is strong, however, or when the
number of jump operators grows extensively with system size (e.g.~for
single-spin decay), these Monte Carlo methods can take a prohibitively
long time to converge, as simulations become dominated by incoherent
jumps that generate large numbers of distinct trajectories that need
to be averaged.

In this work, we develop a new method to simulate short-time dynamics
of collective spin systems with permutationally-symmetric decoherence.
This method is based on a short-time expansion of exact solutions to
the equations of motion for Heisenberg operators.  Evaluating this
expansion requires use of the structure factors of a collective spin
operator algebra; the calculation of these structure factors is one of
the main technical results of this work, which we hope will enable
future studies to develop new analytical insights into the structure
and dynamics of collective spin systems.

@@

In this section we provide the basic theory for our method to compute
expectation values of collective spin operators, deferring lengthy
derivations to the Appendices.  We consider systems composed of $N$
distinct spin-1/2 particles.  Defining individual spin-1/2 operators
$s_{n=\x,\y,\z}\equiv\sigma_n/2$ and
$s_\pm\equiv s_\x\pm is_\y=\sigma_\pm$ with Pauli operators
$\sigma_n$, we denote an operator which acts with $s_n$ on the spin
indexed by $j$ and trivially (i.e.~with the identity $\1$) on all
other spins by $s_n^{(j)}$.  We then define the collective spin
operators $S_n\equiv\sum_{j=1}^Ns_n^{(j)}$ for
$n\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\S_{\v m}}$ as a
basis for all collective spin operators, with
$\v m\equiv\p{m_+,m_\z,m_-}\in\N_0^3$ and
$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$, we can expand all
collective spin Hamiltonians in the form
\begin{align}
  H = \sum_{\v m} h_{\v m} \S_{\v m},
  \label{eq:H_general}
\end{align}
where Hermiticity requires that $h_{\v m}^*=h_{\v m^*}$ for
$\v m^*\equiv\p{m_-,m_\z,m_+}$.  The time evolution of a particular
Heisenberg operator $\S_{\v m}$ under a Hamiltonian of the form in
Eq.~\eqref{eq:H_general} is then given by
\begin{align}
  \f{d}{dt}\S_{\v m}
  &= i\sum_{\v n}h_{\v n}\sp{\S_{\v n},\S_{\v m}}_-
  + \sum_\J \gamma_\J \D\p{\J;\S_{\v m}} \\
  &= \sum_{\v n} T_{\v m\v n} \S_{\v n},
  \label{eq:EOM}
\end{align}
where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\J$ is a set of jump operators
with a corresponding decoherence rate $\gamma_\J$;
\begin{align}
  \D\p{\J;\O}
  \equiv \sum_{J\in\J} \p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+}
\end{align}
is the Heisenberg picture dissipator, or Lindblad superoperator, for
all jump operators $J\in\J$; and $T_{\v m\v n}$ is a matrix element of
the time derivative operator.  Decoherence via uncorrelated decay of
individual spins at a rate $\gamma_-$, for example, would be described
by the set of jump operators
$\J_-=\set{\sigma_-^{(j)}:j=1,2,\cdots,N}$ with
$\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:EOM} can be
determined from the structure factors of $f_{\v n\v m\v\ell}$ of the
general product
$\S_{\v n}\S_{\v m}=\sum_{\v\ell}f_{\v n\v m\v\ell}\S_{\v\ell}$ worked
out in Appendices \ref{sec:identities}--\ref{sec:prod_general}, while
the effects of decoherence from jump operators (i.e.~elements of $\J$)
of the form $\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and
$\Gamma=\sum_n\Gamma_nS_n$ are respectively worked out in Appendices
\ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  These
ingredients are sufficient to compute matrix elements $T_{\v m\v n}$
of the time derivative operator $d/dt$ in Eq.~\eqref{eq:EOM} in most
cases of experimental relevance.

@@

We can expand Heisenberg operators in a Taylor series about $t=0$ to
write
\begin{align}
  \bk{\S_{\v m}}
  = \sum_{k\ge0} \f{t^k}{k!} \bk{\f{d^k}{dt^k} \S_{\v m}}_{t=0}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{\v n} T^k_{\v m\v n} \bk{\S_{\v n}}_{t=0},
  \label{eq:time_series}
\end{align}
where the matrix elements $T^k_{\v m\v n}$ of the $k$-th time
derivative are
\begin{align}
  T^0_{\v m\v n} &\equiv \delta_{\v m\v n}, \\
  T^1_{\v m\v n} &\equiv T_{\v m\v n}, \\
  T^{k>1}_{\v m\v n}
  &\equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
  T_{\v m\v p_1} T_{\v p_1\v p_2} T_{\v p_2\v p_3}
  \cdots T_{\v p_{k-1}\v n},
\end{align}
with $\delta_{\v m\v n}=1$ if $\v m=\v n$ and 0 otherwise.  For
sufficiently short times, we can truncate the series in
Eq.~\eqref{eq:time_series} by taking
\begin{align}
  \bk{\S_{\v m}}
  \to \sum_{k=0}^M \f{t^k}{k!}
  \sum_{\v n} T^k_{\v m\v n} \bk{\S_{\v n}}_{t=0}.
  \label{eq:order_truncation}
\end{align}
Unlike the truncation in Eq.~\eqref{eq:weight_truncation}, the nonzero
matrix elements $T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ now tell us
which operators are relevant for computing the expectation value
$\bk{\S_{\v m}}$ to a fixed order $M$.  Note that the relation
$\S_{\v m^*}=\S_{\v m}^\dag$, which by Hermitian conjugation of
Eq.~\eqref{eq:EOM} also implies that
$T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can be exploited to cut both the
number of expectation values $\bk{\S_{\v m}}_{t=0}$ and matrix
elements $T_{\v m\v n}$ that we need to explicitly compute roughly in
half.

In principle, the factorial suppression of terms at higher orders
(i.e.~in $k$) of the expansion in Eq.~\eqref{eq:order_truncation}
implies for any given time $t$, there exists a truncation order $M_t$
for which truncation error is negligibly small.  In practice, only a
fixed maximal truncation order $\overline M$ is accessible with
limited computational resources, such that
Eq.~\eqref{eq:order_truncation} only allows computation of the
correlator $\bk{\S_{\v m}}$ to a maximal time $t_{\overline M}$.  The
maximal accessible truncation order $\overline M$ generally depends on
the choice of operator basis for collective spin systems, which we
discuss briefly at the end of Appendix \ref{sec:new_basis}.  As we
will see in the following sections, our choice of basis elements
$\S_{\v m}=S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ allows us to accurately
compute spin correlators to times that are relevant e.g.~for spin
squeezing protocols.

@@

To benchmark our method for computing collective spin correlators, we
consider three collective spin models known to dynamically generate
spin-squeezed states: the one-axis twisting (OAT), two-axis twisting
(TAT), and twist-and-turn (TNT) models described by the collective
spin Hamiltonians\cite{kitagawa1993squeezed, micheli2003manyparticle}
\begin{align}
  \label{eq:OAT}
  H_{\t{OAT}} &= \chi S_\z^2, \\
  \label{eq:TAT}
  H_{\t{TAT}} &= \f{\chi}{3} \p{S_\z^2 - S_\y^2}, \\
  \label{eq:TNT}
  H_{\t{TNT}} &= \chi S_\z^2 + \Omega S_\x,
\end{align}
where we include a factor of $1/3$ in the TAT Hamiltonian because it
naturally appears in realistic proposals to experimentally realize
TAT\cite{liu2011spin, huang2015twoaxis}.  For simplicity, we further
fix $\Omega=S\chi$ (with $S\equiv N/2$ throughout this work) to the
critical value known to maximize the entanglement generation rate of
TNT in the large-$N$ limit\cite{micheli2003manyparticle,
  sorelli2015fast}.

@@

The Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} squeeze the
initial product state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
with $S_\x\ket{\X}=S\ket{\X}$.  Our measure of spin squeezing is the
directionally-unbiased Ramsey squeezing parameter determined by the
maximal gain in resolution $\Delta\phi$ of a phase angle $\phi$ over
that achieved by any spin-polarized product state
(e.g.~$\ket{\X}$)\cite{wineland1992spin, ma2011quantum},
\begin{align}
  \xi^2
  \equiv \f{\p{\Delta\phi_{\t{min}}}^2}{\p{\Delta\phi_{\t{polarized}}}^2}
  = \f{N}{\abs{\bk{\v S}}^2}
  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
  \label{eq:squeezing}
\end{align}
where $\v S\equiv\p{S_\z,S_\x,S_\y}$ is a collective spin vector, and
the minimization is performed over all unit vectors $\uv n$ in the
plane orthogonal to the mean spin vector $\bk{\v S}$.  This squeezing
parameter is entirely determined by one- and two-spin correlators of
the form $\bk{S_m}$ and $\bk{S_m S_n}$.  In the case of coherent
dynamics discussed here, these correlators are obtainable via exact
simulations of quantum dynamics in the $\p{N+1}$-dimensional Dicke
manifold of states $\set{\ket{S,m}}$ with net spin $S$ and spin
projection $m$ onto the $z$ axis, i.e.~with
$\bk{S,m|\v S^2|S,m}=S\p{S+1}$ and $\bk{S,m|S_\z|S,m}=m$ for
$m\in\set{-S,-S+1,\cdots,S}$.  In the presence of
permutationally-symmetric decoherence, these correlators are
obtainable with the collective-spin quantum trajectory Monte Carlo
method recently developed in ref.~\cite{zhang2018montecarlo}.  In this
work, these exact and quantum trajectory simulations will be used to
benchmark the expansion in Eq.~\eqref{eq:order_truncation}.

\begin{figure}
  \centering
  \includegraphics{coherent.pdf}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
    under coherent evolution, as computed using exact methods (solid
    lines) and the expansion in Eq.~\eqref{eq:order_truncation} with
    $M=40$ (dashed lines).  Solid circles mark the times at which the
    latter method gives an unphysical result with $\xi^2<0$.  Both
    methods agree exactly in the OAT case for all times shown.}
  \label{fig:coherent}
\end{figure}

\begin{figure}
  \centering
  % \includegraphics{weak_decoherence.pdf}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$ with
    decoherence via uncorrelated decay, excitation, and dephasing of
    individual spins at rates $\gamma_-=\gamma_+=\gamma_\z=\chi$.
    Squeezing is computed using an exact simulation (solid line),
    quantum trajectory methods (dots), and the expansion in
    Eq.~\eqref{eq:order_truncation} with $M=40$ (dashed lines).  Solid
    circles mark the times at which the latter method gives an
    unphysical result with $\xi^2<0$.}
  \label{fig:weak_decoherence}
\end{figure}

>>>>

%%% figures
\usepackage{graphicx} % for figures
\usepackage[caption=false]{subfig} % subfigures (via \subfloat[]{})
\newcommand{\sref}[1]{\protect\subref{#1}} % for referencing subfigures
\graphicspath{{./figures/}} % set path for all figures

@@

\title{Structure constants and short-time simulations of open
  collective quantum spin systems}

@@

Collective spin systems are a versatile resource in quantum science
for a range of applications including quantum-enhanced metrology,
quantum simulation, and quantum information.  Theoretical interest in
such systems dates back to the mid-twentieth century with the
introduction of the Dicke model\cite{dicke1954coherence} that
describes atoms cooperatively interacting with a single mode of the
radiation field and the Lipkin-Meshkov-Glick (LMG) model, a toy model
for testing many-body approximation methods in contemporary nuclear
physics\cite{lipkin1965validity, meshkov1965validity,
  glick1965validity}.  On the experimental side, the development of
advanced trapping, cooling, and control techniques in atomic,
molecular, and optical (AMO) systems have enabled the realization of
collective spin models in a broad range of platforms, including atomic
ensembles\cite{takano2009spin, appel2009mesoscopic, chen2010heralded,
  schleier-smith2010states, chen2011conditional}, Bose-Einstein
condensates\cite{riedel2010atomchipbased, gross2010nonlinear,
  klinder2015dynamical}, ultracold Fermi
gasses\cite{martin2013quantum, smale2018observation}, trapped
ions\cite{bohnet2016quantum}, optical
cavities\cite{leroux2010implementation, bohnet2014reduced,
  cox2016deterministic, hosten2016measurement, hosten2016quantum,
  zhiqiang2017nonequilibrium, lewis-swan2018robust,
  norcia2018cavitymediated}, and molecular
sensors\cite{jones2009magnetic}.  These implementations have triggered
studies of a variety of rich subjects, including quantum
criticality\cite{latorre2005entanglement, alcalde2007functional,
  morrison2008dissipationdriven, morrison2008collective,
  sarandy2009classical, wang2012quantum, kessler2012dissipative,
  majd2014lmg}, non-equilibrium
phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
  bhattacherjee2014nonequilibrium, klinder2015dynamical,
  maghrebi2016nonequilibrium, zhiqiang2017nonequilibrium,
  lang2018concurrence, smale2018observation}, and precision
metrology\cite{wineland1992spin, itano1993quantum,
  kitagawa1993squeezed, ma2011quantum, takano2009spin,
  appel2009mesoscopic, chen2010heralded, schleier-smith2010states,
  chen2011conditional, riedel2010atomchipbased, gross2010nonlinear,
  zhong2010simplified, agarwal1997atomic, lau2014proposal,
  huang2015quantum, wineland1992spin, kitagawa1993squeezed,
  schleier-smith2010squeezing, leroux2010implementation,
  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
  jones2009magnetic}.

One of the primary motivations for studying collective spin systems is
their application to quantum-enhanced metrology.  Quantum projection
noise limits the error $\Delta\phi$ in the measurement of a phase
angle $\phi$ with $N$ independent spins to
$\Delta\phi\sim1/\sqrt{N}$\cite{wineland1992spin, itano1993quantum,
  ma2011quantum}.  Collective spin systems enable a means to break
through this limit via the preparation of many-body entangled states
such as NOON\cite{jones2009magnetic, chen2010heralded,
  zhong2010simplified}, spin-cat\cite{agarwal1997atomic,
  lau2014proposal, huang2015quantum}, and spin-squeezed
states\cite{wineland1992spin, kitagawa1993squeezed, takano2009spin,
  appel2009mesoscopic, schleier-smith2010states,
  schleier-smith2010squeezing, leroux2010implementation,
  riedel2010atomchipbased, gross2010nonlinear, chen2011conditional,
  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
  ma2011quantum} that allow for measurement errors
$\Delta\phi\sim1/N^\epsilon$ with $\epsilon>1/2$.  Generating such
states typically requires nonlinear dynamics that are realized using
collisional, photon-mediated, or phonon-mediated interactions.
Although a truly collective spin model requires uniform, all-to-all
interactions, as long as measurements do not distinguish between
constituent particles even non-uniform systems can be effectively
described by a uniform model with renormalized
parameters\cite{hu2015entangled}.

In the absence of decoherence, permutation symmetry and total spin
conservation divide the total Hilbert space of a collective spin
system into superseletion sectors that grow only linearly with system
size $N$, thereby admitting efficient classical simulation of its
dynamics.  Decoherence will generally violate total spin conservation
and require the use of density operators, increasing the dimension of
accessible state space to $\sim N^3$\cite{hartmann2016generalized,
  xu2013simulating} and allowing for exact simulations of
$N\lesssim100$ particles.  If decoherence is sufficiently weak,
dynamics can be numerically solvable for $N\lesssim10^5$ particles via
``quantum trajectory'' Monte Carlo methods\cite{plenio1998quantumjump,
  zhang2018montecarlo} (also known as ``quantum jump'' or ``Monte
Carlo wavefunction'' methods) that can reproduce all expectation
values of interest.  When decoherence is strong, however, or when the
number of jump operators grows extensively with system size (e.g.~for
single-spin decay), these Monte Carlo methods can take a prohibitively
long time to converge, as simulations become dominated by incoherent
jumps that generate large numbers of distinct quantum trajectories
that need to be averaged to compute expectation values.

In this work, we develop a new method to simulate short-time dynamics
of collective spin systems with permutationally-symmetric decoherence.
This method is based on a short-time expansion of exact solutions to
the equations of motion for Heisenberg operators.  Evaluating this
expansion requires using the structure constants of a collective spin
operator algebra; the calculation of these structure constants is one
of the main technical results of this work, which we hope will enable
future studies to develop new analytical insights into the dynamics of
collective spin systems.

@@

In this section we provide the basic theory for our method to compute
expectation values of collective spin operators, deferring lengthy
derivations to the Appendices.  We consider systems composed of $N$
distinct spin-1/2 particles.  Defining individual spin-1/2 operators
$s_{n=\x,\y,\z}\equiv\sigma_n/2$ and
$s_\pm\equiv s_\x\pm is_\y=\sigma_\pm$ with Pauli operators
$\sigma_n$, we denote an operator that acts with $s_n$ on the spin
indexed by $j$ and trivially (i.e.~with the identity $\1$) on all
other spins by $s_n^{(j)}$.  We then define the collective spin
operators $S_n\equiv\sum_{j=1}^Ns_n^{(j)}$ for
$n\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\S_{\v m}}$ as a
basis for all collective spin operators, with
$\v m\equiv\p{m_+,m_\z,m_-}\in\N_0^3$ and
$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$, we can expand all
collective spin Hamiltonians in the form
\begin{align}
  H = \sum_{\v m} h_{\v m} \S_{\v m},
  \label{eq:H_general}
\end{align}
where Hermiticity requires that $h_{\v m}^*=h_{\v m^*}$ for
$\v m^*\equiv\p{m_-,m_\z,m_+}$.  The time evolution of a particular
Heisenberg operator $\S_{\v m}$ under a Hamiltonian of the form in
Eq.~\eqref{eq:H_general} is then given by
\begin{align}
  \f{d}{dt}\S_{\v m}
  &= i\sum_{\v n}h_{\v n}\sp{\S_{\v n},\S_{\v m}}_-
  + \sum_\J \gamma_\J \D\p{\J;\S_{\v m}} \\
  &= \sum_{\v n} T_{\v m\v n} \S_{\v n},
  \label{eq:EOM}
\end{align}
where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\J$ is a set of jump operators
with a corresponding decoherence rate $\gamma_\J$;
\begin{align}
  \D\p{\J;\O}
  \equiv \sum_{J\in\J} \p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+}
\end{align}
is the Heisenberg picture dissipator, or Lindblad superoperator, for
all jump operators $J\in\J$; and $T_{\v m\v n}$ is a matrix element of
the time derivative operator.  Decoherence via uncorrelated decay of
individual spins at a rate $\gamma_-$, for example, would be described
by the set of jump operators
$\J_-=\set{\sigma_-^{(j)}:j=1,2,\cdots,N}$ with
$\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:EOM} can be
determined from the structure constants $f_{\v n\v m\v\ell}$ of the
general product
$\S_{\v n}\S_{\v m}=\sum_{\v\ell}f_{\v n\v m\v\ell}\S_{\v\ell}$ worked
out in Appendices \ref{sec:identities}--\ref{sec:prod_general}, while
the effects of decoherence from jump operators (i.e.~elements of $\J$)
of the form $\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and
$\Gamma=\sum_n\Gamma_nS_n$ are respectively worked out in Appendices
\ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  These
ingredients are sufficient to compute matrix elements $T_{\v m\v n}$
of the time derivative operator $d/dt$ in Eq.~\eqref{eq:EOM} in most
cases of experimental relevance.

@@

We can expand Heisenberg operators in a Taylor series about $t=0$ to
write
\begin{align}
  \bk{\S_{\v m}}
  = \sum_{k\ge0} \f{t^k}{k!} \bk{\f{d^k}{dt^k} \S_{\v m}}_{t=0}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{\v n} T^k_{\v m\v n} \bk{\S_{\v n}}_{t=0},
  \label{eq:time_series}
\end{align}
where the matrix elements $T^k_{\v m\v n}$ of the $k$-th time
derivative are
\begin{align}
  T^0_{\v m\v n} &\equiv \delta_{\v m\v n}, \\
  T^1_{\v m\v n} &\equiv T_{\v m\v n}, \\
  T^{k>1}_{\v m\v n}
  &\equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
  T_{\v m\v p_1} T_{\v p_1\v p_2} T_{\v p_2\v p_3}
  \cdots T_{\v p_{k-1}\v n},
\end{align}
with $\delta_{\v m\v n}=1$ if $\v m=\v n$ and 0 otherwise.  For
sufficiently short times, we can truncate the series in
Eq.~\eqref{eq:time_series} by taking
\begin{align}
  \bk{\S_{\v m}}
  \to \sum_{k=0}^M \f{t^k}{k!}
  \sum_{\v n} T^k_{\v m\v n} \bk{\S_{\v n}}_{t=0}.
  \label{eq:order_truncation}
\end{align}
Unlike the truncation in Eq.~\eqref{eq:weight_truncation}, the nonzero
matrix elements $T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ now tell us
which operators are relevant for computing the expectation value
$\bk{\S_{\v m}}$ to a fixed order $M$.  Note that the relation
$\S_{\v m^*}=\S_{\v m}^\dag$, which by Hermitian conjugation of
Eq.~\eqref{eq:EOM} also implies that
$T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can be exploited to cut both the
number of expectation values $\bk{\S_{\v m}}_{t=0}$ and matrix
elements $T_{\v m\v n}$ that we may need to explicitly compute roughly
in half.

In principle, the factorial suppression of terms at higher orders
(i.e.~in $k$) of the expansion in Eq.~\eqref{eq:order_truncation}
implies for any given time $t$, there exists a truncation order $M_t$
for which truncation error is negligibly small.  In practice, only a
fixed maximal truncation order $M_{\t{max}}$ is accessible with
limited computational resources, such that
Eq.~\eqref{eq:order_truncation} only allows computation of the
correlator $\bk{\S_{\v m}}$ to a maximal time $t_{M_{\t{max}}}$.  The
maximal accessible truncation order $M_{\t{max}}$ generally depends on
the choice of operator basis for collective spin systems, which we
discuss briefly in Section \ref{sec:breakdown} and Appendix
\ref{sec:new_basis}.  As we will see in the following sections, our
choice of basis elements $\S_{\v m}=S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$
allows us to accurately compute spin correlators to times that are
relevant e.g.~for spin squeezing protocols.

@@

To benchmark our method for computing collective spin correlators, we
consider three collective spin models known to dynamically generate
spin-squeezed states: the one-axis twisting (OAT), two-axis twisting
(TAT), and twist-and-turn (TNT) models described by the collective
spin Hamiltonians\cite{kitagawa1993squeezed, micheli2003manyparticle}
\begin{align}
  \label{eq:OAT}
  H_{\t{OAT}} &= \chi S_\z^2, \\
  \label{eq:TAT}
  H_{\t{TAT}} &= \f{\chi}{3} \p{S_\z^2 - S_\y^2}, \\
  \label{eq:TNT}
  H_{\t{TNT}} &= \chi S_\z^2 + \Omega S_\x,
\end{align}
where we include a factor of $1/3$ in the TAT Hamiltonian because it
naturally appears in realistic proposals to experimentally implement
TAT\cite{liu2011spin, huang2015twoaxis}.  For simplicity, we further
fix $\Omega=S\chi$ (with $S\equiv N/2$ throughout this work) to the
critical value known to maximize the entanglement generation rate of
TNT in the large-$N$ limit\cite{micheli2003manyparticle,
  sorelli2015fast}.

@@

The Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} squeeze the
initial product state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
with $S_\x\ket{\X}=S\ket{\X}$.  Our measure of spin squeezing is the
directionally-unbiased Ramsey squeezing parameter determined by the
maximal gain in resolution $\Delta\phi$ of a phase angle $\phi$ over
that achieved by any spin-polarized product state
(e.g.~$\ket{\X}$)\cite{wineland1992spin, ma2011quantum},
\begin{align}
  \xi^2
  \equiv \f{\p{\Delta\phi_{\t{min}}}^2}{\p{\Delta\phi_{\t{polarized}}}^2}
  = \f{N}{\abs{\bk{\v S}}^2}
  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
  \label{eq:squeezing}
\end{align}
where $\v S\equiv\p{S_\z,S_\x,S_\y}$ is a collective spin vector, and
the minimization is performed over all unit vectors $\uv n$ in the
plane orthogonal to the mean spin vector $\bk{\v S}$.  This squeezing
parameter is entirely determined by one- and two-spin correlators of
the form $\bk{S_m}$ and $\bk{S_m S_n}$.  In the case of coherent
dynamics discussed here, these correlators are obtainable via exact
simulations of quantum dynamics in the $\p{N+1}$-dimensional Dicke
manifold of states $\set{\ket{S,m}}$ with net spin $S$ and spin
projection $m$ onto the $z$ axis, i.e.~with
$\bk{S,m|\v S^2|S,m}=S\p{S+1}$ and $\bk{S,m|S_\z|S,m}=m$ for
$m\in\set{-S,-S+1,\cdots,S}$.  In the presence of
permutationally-symmetric decoherence, these correlators are
obtainable with the collective-spin quantum trajectory Monte Carlo
method developed in ref.~\cite{zhang2018montecarlo}.  In this work,
these exact and quantum trajectory simulations will be used to
benchmark the expansion in Eq.~\eqref{eq:order_truncation}.

\begin{figure}
  \centering
  \subfloat[Squeezing under coherent dynamics \label{fig:coherent}]
  {\includegraphics{coherent.pdf}} \\
  \subfloat[Squeezing with decoherence:
  $\gamma_-=\gamma_+=\gamma_\z=\chi$ \label{fig:incoherent}]
  {\includegraphics{decoherence_weak.pdf}}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
    under \sref{fig:coherent} coherent and \sref{fig:incoherent}
    incoherent evolution, as computed using exact methods (solid
    lines), quantum trajectory simulations (dots), and the expansion
    in Eq.~\eqref{eq:order_truncation} with $M=40$ (darker dashed
    lines).  Solid circles mark the times at which the last of these
    methods gives an unphysical result with $\xi^2<0$.}
  \label{fig:benchmarking}
\end{figure}

Figure \ref{fig:benchmarking} compares the squeezing parameter $\xi^2$
for $N=10^4$ spins initially in the state $\ket{\X}$ evolved under the
Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}, as computed via
both benchmarking simulations and the prescription in
Eq.~\eqref{eq:order_truncation} with $M=40$.  This squeezing is shown
for both coherent dynamics (Figure \ref{fig:coherent}), as well as
dynamics in the presence of uncorrelated emission, absorption, and
dephasing of individual spins at a rate $\chi$ (Figure
\ref{fig:incoherent}), respectively described by the sets of jump
operators $\J_n=\set{s_n^{(j)}}$ with corresponding decoherence rates
$\gamma_n=\chi$ for $n\in\set{-,+,\z}$.  The squeezing results shown
in Figure \ref{fig:benchmarking} were computed in a rotated basis with
$\p{s_\z,s_\x}\to\p{s_\x,-s_\z}$ and
$\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$; the only effect of this
rotation on the results presented in this figure is to roughly double
the time for which all TNT results agree.  The reason for different
results in rotated bases has to do with the breakdown of the expansion
in Eq.~\eqref{eq:order_truncation}, which we discuss further in
Section \ref{sec:breakdown}.
